---
title: "p8105_hw3_xf2226"
author: "Xin Fang"
date: "10/19/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(p8105.datasets)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  digits = 3
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d 

knitr::opts_chunk$set(
  fig.height = 6,
  fig.width = 8,
  message = F,
  warning = F 
  )
```

# Problem 1 
```{r}
data("instacart")

instacart <- instacart %>% 
    mutate(
      order_dow = order_dow + 1,
      order_dow = wday(order_dow, label = T)
   ) 
```
## A short description of the dataset, noting the size and structure of the data, describing some key variables
This instacart dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

The observations are orders which is included in the `order_id` column. We have `r length(unique(pull(instacart, order_id)))` order in total. Within those orders, there are `r length(unique(pull(instacart, product_name)))` unique products. Some examples of the products include `r head(unique(pull(instacart, product_name)),5)`. Those products are coming from `r length(unique(pull(instacart, aisle_id)))` unique aisles and `r length(unique(pull(instacart, department)))` departments. The top two days of the week where the most orders are placed is `r tail(names(sort(table(pull(instacart,order_dow)))), 2)`, which are the weekends. 

Some key varaibles are: 

* `order_id`= order identifier
* `product_id`= product identifier
* `user_id`= customer identifier
* `order_number`= the order sequence number for this user (1=first, n=nth)
* `order_dow`= the day of the week on which the order was placed
* `product_name`= name of the product
* `aisle_id`= aisle identifier
* `aisle`= the name of the aisle
* `department_id`= department identifier
* `department`= the name of the department

## Illstrative examples of observations
```{r}
instacart_dow_dept <- instacart %>% 
  filter(department == "produce" | department == "frozen") %>% 
  group_by(order_dow, aisle, department) %>% 
  summarise(total = n())

instacart_dow_dept %>% 
  ggplot(aes(x = order_dow, y = total, group = aisle, color = aisle)) + 
  geom_line() +
  facet_grid(department ~ ., scales = "free_y") +
   labs(
    x = "Day of the Week",
    y = "Quantity of Orders",
  )
```

## Problem 1.1 
How many aisles, and which are most items from?
```{r}
aisles <- instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```
There are `r nrow(aisles)` number of aisles. The top 3 aisles where most items are ordered are from `r head(pull(aisles, aisle), 3)`

## Problem 1.2
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
aisles %>%
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + geom_point() +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1)) + 
  labs(
    y = "number of items ordered", 
    title = "Number of items ordered from each aisle (that has more than 10000 items ordered)"
  )
```


## Problem 1.3 
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% #top three popular items
  arrange(aisle, rank) %>% 
  knitr::kable()
```


## Problem 1.4
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% #taking mean hour of the day
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```

# Problem 2 
## load and clean BRFSS dataset
```{r load and clean BRFSS dataset}
data("brfss_smart2010")
brfss_smart2010 <- brfss_smart2010 %>% 
  janitor::clean_names() 

brfss <- brfss_smart2010 %>%  
  filter(topic %in% "Overall Health") %>%
  filter(response %in% c("Poor","Fair","Good","Very good","Excellent")) %>%
  mutate(response = factor(response, ordered = TRUE, 
                       levels = c("Poor", "Fair", "Good", "Very good", "Excellent")),
         locationdesc = gsub('^.{0,5}', '', locationdesc)) %>% 
  rename(state = locationabbr, county = locationdesc)
```

## Problem 2.1: 
In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
states_2002 <- brfss %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarise(n_state = n()) %>% 
  filter(n_state > 6)

states_2010 <- brfss %>% 
  filter(year == 2010) %>% 
  group_by(state) %>% 
  summarise(n_state = n()) %>% 
  filter(n_state > 6)
```
In 2002, the states that were observed at 7 or more locations are `r pull(states_2002, state)`. \
In 2010, the states that were observed at 7 or more locations are `r pull(states_2010, state)`.

## Problem 2.2 
Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
brfss %>% 
  filter(response == "Excellent") %>% 
  group_by(state, year) %>%  
  mutate(
    data_value_ave = mean(data_value, na.rm = TRUE)
  ) %>% 
  select(year, state, data_value_ave) %>% 
  unique() %>% 
  ggplot(
    aes(x = year, y = data_value_ave, group = state, color = state)) + 
  geom_line() + 
  labs(
    x = "Average Data Value"
  )
```


## Problem 2.3 
Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss %>% 
  filter(year == "2006" | year == "2010", state == "NY") %>% 
  select(year, data_value, response) %>%
  ggplot(
    aes(x = data_value, color = response)) + 
  geom_density() + 
  facet_grid(.~ year) +
  labs(
    x = "Data Value"
  )
```

# Problem 3 
## load and tidy accelerometer dataset
```{r load and tidy accelerometer dataset}
accel <- read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() 

accel_cleaned <- accel %>% 
  pivot_longer(
    cols = activity_1:activity_1440,
    names_to = "activity_min", 
    values_to = "activity_count"
  ) %>% 
  mutate(
    activity_min = gsub("^.{0,9}", "", activity_min), 
    day = factor(day, ordered = TRUE, 
                  levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")), 
    activity_min = as.numeric(activity_min), 
    day_of_week = as.factor(ifelse(day %in% c("Saturday", "Sunday"), "weekend", "weekday"))
  ) %>% 
  select(week, day_id, day, day_of_week, activity_min, activity_count)
```
This cleaned accelerometer dataset contains `r nrow(accel_cleaned)` observations and `r ncol(accel_cleaned)` variables. 

Some key varaibles are: 

* `week`= week of the observation
* `day_id`= day of the observation
* `day`= day of the week of the observation
* `activity_min`= min of the observation
* `activity_count`= activity counts for each minute
* `day_of_week`= what is weekday vs. weekend


## Problem 3.1 
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
accel_cleaned %>% 
  group_by(week, day) %>% 
  summarize(total_activity = sum(activity_count)) %>% 
  pivot_wider(
    names_from = day,
    values_from = total_activity
  ) %>% 
  knitr::kable(digits = 1)
```

I did not see any apparent trend. However, Saturday for week 4 and 5 have significantly less activity count compared to any other days of the study period. We need to look more in-depth about those two days and see if some serious measurement errors occurred 

## Problem 3.2 
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r, warning=FALSE}
accel_cleaned %>%
  mutate(hour = activity_min/60) %>%
  ggplot(aes(x = hour, y = activity_count, group = day_id)) +
  geom_point(aes(color = day)) +
  geom_smooth(aes(color = day, group = day))  +
  scale_x_continuous(breaks = c(0, 5, 10, 15, 20, 24)) +
  labs(
    x = "Hour of the day",
    y = "Accelerometer activity count", 
    title = '24-hour activity time courses for each day'
  ) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1), plot.title = element_text(size = 11))
```



